Optimal strategies in mental rotation tasks

### 2013-05-13 man/final-project.tex
Clean up implementation section, up to gold standard model

We define the prior probabilities over stimuli based on how they are
generated. For shapes with $n$ vertices, each vertex is at a random
angle with a radius random chosen between 0 and 1, in polar
coordinates, and could be chosen in any of $n!$ different ways. Thus,
the prior over shapes is:
\begin{equation}
\subsection{Likelihood when the configurations are distinct}

  p(X)=n!\left(\frac{1}{2\pi}\right)^n
\end{equation} 
And so the joint likelihood is just:
\begin{equation}
  p(X_a, X_b\ \vert \ h_0)=\left[n!\left(\frac{1}{2\pi}\right)^n\right]^2
\end{equation}

\subsection{Likelihood when the configurations are equivalent}

We cannot know the quantity $p(X_b\vert X_R)$ exactly, so we
approximate it with a similarity function $S(X_b, X_R)$, which also
takes into account the different possible mappings of vertices. There
are $n$ mappings $M$ of the $n$ vertices: because some vertices are
connected to form a closed loop, we assume the uncertainty is only in
which is the ``first'' vertex. So, the possible orderings are of the
form $M=\lbrace{}0, 1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0,
\ldots{}, n-1\rbrace{}$, and so on. This gives us the final form of
the similarity function:
\begin{equation}
  S(X_b, X_R)=\frac{1}{n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\ \vert \ (MX_R)[i], \Sigma)
\end{equation}
Where $i$ denotes the $i^{th}$ vertex. We do not know the closed,
analytic form of this function, but we can evaluate it for any $X_b$
and $X_R$. Given this, the desired quantity for $h_1$ becomes:
\begin{equation}
  Z = p(X_a,X_b\ \vert \ h_1)\approx \int_R S(X_b, X_R)p(X_a)p(R)\ \mathrm{d}R
\end{equation}
Note that for brevity, from here on we will refer to $p(X_a, X_b\
\vert \ h_1)$ as $Z$.

Finally, computing any single $X_R$ is a costly operation in that it
is both a lengthy operation and memory-intensive, so we want to
minimize the number of $X_R$ that are computed. With this in mind, we
now examine several different approaches to estimating $S$ and,
ultimately, $Z$.

>>>>

We define the prior probabilities over stimuli based on how they are
generated. For shapes with $n$ vertices, each vertex is at a random
angle with a radius random chosen between 0 and 1, in polar
coordinates, and could be chosen in any of $n!$ different ways. Thus,
the prior over a shape $X$ is:
\begin{equation}
  p(X)=n!\left(\frac{1}{2\pi}\right)^n
  \label{eq:prior}
\end{equation} 
which gives us the denominator in Equation \ref{eq:lh-ratio}.

Computing the numerator of Equation \ref{eq:lh-ratio} is more
difficult, as we do know $p(X_b\vert X_R)$. We approximate it with a
similarity function $S(X_b, X_R)$, which also takes into account the
different possible mappings of vertices:
\begin{equation}
  Z=\int_R S(X_b, X_R)p(R)\ \mathrm{d}R\ \approx\int_Rp(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R
  \label{eq:Z}
\end{equation}

Because the vertices are connected in a way which forms a closed loop,
we need only consider $n$ mappings, $M$, of the $n$ vertices (we
assume the uncertainty is only in which is the ``first''
vertex). Thus, the possible orderings are of the form $M=\lbrace{}0,
1, \ldots{}, n\rbrace{}$, $M=\lbrace{}n, 0, \ldots{}, n-1\rbrace{}$,
and so on. This gives us an explicit form for the similarity function:
\begin{equation}
  S(X_b, X_R)=\frac{1}{n}\sum_{M\in\mathbb{M}}\prod_{i=1}^n\mathcal{N}(X_b[i]\ \vert \ (MX_R)[i], \Sigma)
  \label{eq:similarity}
\end{equation}
where $i$ denotes the $i^{th}$ vertex. An example of $S$ for the
stimuli shown in Figure \ref{fig:stimuli} is illustrated in Figure
\ref{fig:likelihood}.

This process of mental rotation (i.e., generating a single $X_R$ (as
in Equation \ref{eq:tau}) and then computing $S(X_b, X_R)$) is a
computationally demanding cognitive process. Thus, our goal is to
minimize the number of rotations while still obtaining an estimate of
$Z$ that is accurate enough to choose the correct hypothesis. With
this in mind, we now examine several approaches to estimating $Z$.
